Codebook for Activity_Tidy.txt
==============================

Human Activity Recognition Using Smartphones Data Set 

Author: Thomas Berger (https://github.com/bergertom/GetDataProject.git)

This codebook was generated on `r as.character(Sys.time())` by running the script ``run_analysis.R``.

**The Codebook Dimensions are described at the end of this document.**

Steps to create
===============

1. Merges the training and the test sets to create one data set
---------------------------------------------------------------
```{r}
# read the column names (used for both train and test raw data) from the file provided
column_names <- read.table(file.path(getwd(), datadir,"features.txt"), col.names=c("index", "column_name"))

# read the test data sets
subject1  <- read.table(file.path(getwd(), datadir, "test" , "subject_test.txt"), col.names=c("Subject"))
data1     <- read.table(file.path(getwd(), datadir, "test" , "X_test.txt"))
activity1 <- read.table(file.path(getwd(), datadir, "test" , "Y_test.txt"), col.names=c("Activity"))
# set column names for data1
colnames(data1) <- column_names[,2]
# column merge test data
test_data <- cbind(subject1,activity1,data1)

# read the data sets
subject2  <- read.table(file.path(getwd(), datadir, "train" , "subject_train.txt"), col.names=c("Subject"))
data2     <- read.table(file.path(getwd(), datadir, "train" , "X_train.txt"))
activity2 <- read.table(file.path(getwd(), datadir, "train" , "Y_train.txt"), col.names=c("Activity"))
# set column names for data2
colnames(data2) <- column_names[,2]
# column merge all data
train_data <- cbind(subject2,activity2,data2)

# row merge to have the final data frame
data <- rbind(test_data, train_data)
```

2. Extracts only the measurements on the mean and standard deviation for each measurement
-----------------------------------------------------------------------------------------
```{r}
# filter for column names coainting 'mean()' or 'std()'
activity_ok <- grepl('mean\\(\\)|std\\(\\)',column_names[,2])

# add first to columns(Subject,Activity); they are not filtered
column_filter <- append (c(TRUE,TRUE), activity_ok)

# filter the columns on the mean and standard deviation 
data_filtered <- data[column_filter]
```

3. Uses descriptive activity names to name the activities in the data set
-------------------------------------------------------------------------
```{r}
# read activity names from the provided file
activity_names<-read.table(file.path(getwd(), datadir, "activity_labels.txt"),sep=" ",col.names=c("activityLabel","Activity"))
```

4. Appropriately labels the data set with descriptive activity names
--------------------------------------------------------------------
```{r}
# map the activity code to an Activity_Name (cast because R is too smart, will use factor)
data_filtered$Activity = as.character(activity_names[data_filtered$Activity,2])
```

5. Creates a second, independent tidy data set with the average of each variable for each activity and each subject
---------------------------------------------------------------------------------------------------------------------
```{r}
# average of each variable for each activity and each subject
data_tidy <- data.frame(data_filtered %>% group_by(Subject, Activity) %>% summarise_each(funs(mean)))

# write the dataset to a file
write.table(data_tidy, file="Activity_Tidy.txt", quote = FALSE, sep = "\t", row.names = FALSE)
```

Codebook Dimensions
===================

The data-set ``Activity_Tidy.txt`` has `r ncol(data_tidy)` columns and `r nrow(data_tidy)` rows.

The Subjects are defined in ``X__test.txt`` (rows=`r nrow(subject1)`) and ``Y_train.txt`` (rows=`r nrow(subject2)`).

The Activity code was mapped using the file ``activity_labels.txt``; there are `r nrow(activity_names)` activities:
`r paste(activity_names[,2],collapse=",")`.

The rest of 66 variables are all the average of mean values and standard deviation values for data collected from different sensors.
 
The original data contains `r nrow(column_names) ` different measurement values; which are explained in the given file ``features.txt``.
The data for this database come from combining the files ``Y_test.txt`` and ``Y_train.txt`` files. Then only columns for mean and average are filtered.
The tidy data-set contains the average of each variable for each activity and each subject.

```{r }
dim(data_tidy)
```

```{r }
str(data_tidy)
```
