Codebook for Activity_Tidy.txt
==============================

Human Activity Recognition Using Smartphones Data Set 

Author: Thomas Berger (https://github.com/bergertom/GetDataProject.git)

This codebook was generated on `r as.character(Sys.time())` by running the script ``run_analysis.R``.

**The Codebook Dimensions are described at the end of this document.**

Steps to create
===============

1. Merges the training and the test sets to create one data set
---------------------------------------------------------------
```{r}
# read the column names (used for both train and test raw data)
column_names <- read.table(file.path(getwd(), datadir,"features.txt"), col.names=c("index", "column_name"))

# read the test data sets
subject1  <- read.table(file.path(getwd(), datadir, "test" , "subject_test.txt"), col.names=c("Subject"))
data1     <- read.table(file.path(getwd(), datadir, "test" , "X_test.txt"))
activity1 <- read.table(file.path(getwd(), datadir, "test" , "Y_test.txt"), col.names=c("Activity"))
# add column names to data1
colnames(data1) <- column_names[,2]
# column merge test data
test_data <- cbind(subject1,activity1,data1)

# read train data sets
subject2  <- read.table(file.path(getwd(), datadir, "train" , "subject_train.txt"), col.names=c("Subject"))
data2     <- read.table(file.path(getwd(), datadir, "train" , "X_train.txt"))
activity2 <- read.table(file.path(getwd(), datadir, "train" , "Y_train.txt"), col.names=c("Activity"))
# add column names to data2
colnames(data2) <- column_names[,2]
# column merge train data
train_data <- cbind(subject2,activity2,data2)

# row merge the final data frame
data <- rbind(test_data, train_data)
```

2. Extracts only the measurements on the mean and standard deviation for each measurement
-----------------------------------------------------------------------------------------
```{r}
# filter for column names coainting 'mean()' or 'std()'
activity_ok <- grepl('mean\\(\\)|std\\(\\)',column_names[,2])

# add first to columns(Subject,Activity) to boolean vector
column_filter <- append (c(TRUE,TRUE), activity_ok)

# filter the columns
data_filtered <- data[column_filter]
```

3. Uses descriptive activity names to name the activities in the data set
-------------------------------------------------------------------------

continue at step 4

4. Appropriately labels the data set with descriptive activity names
--------------------------------------------------------------------
```{r}
# read activity names
activity_names<-read.table(file.path(getwd(), datadir, "activity_labels.txt"),sep=" ",col.names=c("activityLabel","Activity"))

# map the activity code to an Activity_Name (cast because R is too smart, will use factor)
data_filtered$Activity = as.character(activity_names[data_filtered$Activity,2])
```

5. Creates a second, independent tidy data set with the average of each variable for each activity and each subject
---------------------------------------------------------------------------------------------------------------------
```{r}
# copy Subject and Activity
data_tidy <- data_filtered[, 1:2]

# tBodyAcc = mean (BodyAcc-mean()-X, tBodyAcc-mean()-Y,tBodyAcc-mean()-Z, tBodyAcc-std()-X, tBodyAcc-std()-Y,tBodyAcc-std()-Z)
# fBodyBodyGyroJerkMag = mean (fBodyBodyGyroJerkMag-mean, fBodyBodyGyroJerkMag-std)

# here define the new columns for the tidy data-set
# this list is copied directly from the file "features_info.txt" 
columns_data <- c( "tBodyAcc","tGravityAcc","tBodyAccJerk", "tBodyGyro","tBodyGyroJerk","tBodyAccMag", "tGravityAccMag","tBodyAccJerkMag",
                   "tBodyGyroMag", "tBodyGyroJerkMag","fBodyAcc", "fBodyAccJerk","fBodyGyro","fBodyAccMag","fBodyBodyAccJerkMag",
                   "fBodyBodyGyroMag", "fBodyBodyGyroJerkMag" )

# calculate average for each value and document what is done
col_documentation <- c("The Subjects are assigned numbers from 1 to 30.", "The Activity description")
for (col_name in columns_data) {
    # loop through the column names
    col_numbers <- numeric()
    col_temp <- "mean of ("
    for (col_counter in 1:68) {
        acolumn_name  <- colnames(data_filtered)[col_counter]
        if (unlist(strsplit(acolumn_name, "-"))[1] == col_name) {
            # and add all matching names to the vector
            col_numbers <- c(col_numbers, col_counter)
            col_temp <- paste(col_temp, acolumn_name)
        }
    }
    # calculate the average for the given subset of columns
    data_tidy[[col_name]] <- rowMeans(data_filtered[, col_numbers])
    col_temp <- paste(col_temp, ")")
    col_documentation <- c(col_documentation, col_temp)
}

# write the dataset to a file
write.table(data_tidy, file="Activity_Tidy.txt", quote = FALSE, sep = "\t", row.names = FALSE)
```

Dimensions
==========

The data-set ``Activity_Tidy.txt`` has a total of `r nrow(data_tidy)` rows.

```{r, results='asis', echo=FALSE}
   print(kable(data.frame("column name"=colnames(data_tidy), "explanation"=col_documentation)), type = "html")
```

The Subjects are defined in ``X__test.txt`` (rows=`r nrow(subject1)`) and ``Y_train.txt`` (rows=`r nrow(subject2)`).

The Activity code was mapped using the file ``activity_labels.txt`` (rows=`r nrow(activity_names)`).

The original data contains `r nrow(column_names) ` different measurement values; which are explained in the given file ``features.txt``.
The data for this database come from combining the files ``Y_test.txt`` and ``Y_train.txt`` files. Then only columns for mean and average are filtered. Then all data points (X,Y,Z) are combined using the mean() function to create the `r length(columns_data)`  tidy data-set.

```{r }
dim(data_tidy)
```

```{r }
str(data_tidy)
```
